Age-related changes in plant growth and traits

### 2015-04-30 results.Rmd
Continuing to explore patterns of RA, RGR, growth, repro investment Strong correlations between shifts in growth vs. repro and height (within species) is particularly interesting - "cleaner" RA schedules against height than age or weight
continuing to organize results

# LMA versus age
The plot below is a bulked LMA sample from each site, of leaves from the second most recent cohort

There is a significant increase with age (p=0.002) and many species differ from one another, despite the very few points
Age is only borderline significant (p=0.04) when the seedlings are removed

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LMA~age,data=subset(LMA,branch_age=="mid"),log="xy", xlab="age",ylab="LMA",col=col.spp[as.factor(species)],pch=16)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

# Leaf lifespan with age
Leaf lifespan increases from seedlings (1.4 yrs) to saplings (2.4 years) to adults (>4 years), but there are no differences between the different adult populations
```{r, echo=FALSE,results='hide', message=FALSE}
plot(LL_birth~age,SummaryInd,log="xy", xlab="age",ylab="leaf lifespan (based on leaf births)",col=col.spp[as.factor(species)])
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

@@

```{r, echo=FALSE,results='hide', message=FALSE}
col.LMA <-c("purple", "pink", "light green", "orange","blue", "light blue","red")
labels.LMA <-c("age 1.4", "age 2.4", "age 5", "age 7", "age 9", "age 32", "HATE")
plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=17)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
```

# LMA vs LL without young plants
The pattern becomes more obvious when young plants (and Hakea) are omitted

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LL_bd_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE" & age>1.5), pch=16, cex=1.2, col=col.age2[as.factor(age)])
legend("bottomright",col=col.age2,labels.age2,pch=16)
```

@@

# Species with greater wood density have higher growth rates
Again there is no relationship in younger plants
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==5), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```

# Species with larger seeds have higher growth rates
The same basic relationship exists across all ages, so combined all data here
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ seed_size,data=subset(SummarySppAge,age>.5),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="seed size (mg)",ylab="growth investment (mg per year)",main="Effect of seed size on growth rates")
legend("topleft",col=col.age, labels.age, pch=16, cex=1)
text(GrowthInv_mean ~ seed_size, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

# Species with greater maximum height have higher growth rates
For younger plants there is no relationship with *SPECIES* overall maximum H, but there is a positive correlation with the maximum height the species reaches at a given age -i.e. fast growing young plants are tall when young.
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.2,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

# Correlations between different growth measures
Increased cross-sectional area of a leader and the entire plant are fairly correlated and uneffected by age (once not baby)
```{r, echo=FALSE,results='hide', message=FALSE}
plot(change_shoot_area~change_basal_area,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=.8,xlab="change in basal stem area",ylab="change in shoot area")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1)
```

>>>>

@@

plot(LL_birth_mean ~ LMA_mean, SummarySppAge, type="n", xlab="LMA", ylab="leaf lifespan (years)", main="leaf lifespan versus LMA (all species and ages)", las=1,pch=17, log="xy")
points(LL_birth_mean ~ LMA_mean, data=subset(SummarySppAge, species!="HATE"), pch=16, cex=1.2, col=col.age[as.factor(age)])
points(LL_birth_mean ~ LMA_mean, col="red", data=subset(SummarySppAge, species=="HATE"), pch=17)
legend("bottomright",col=col.LMA, labels.LMA, pch=16, cex=1.4)
```

@@

#22 April 2015
Generic code to create plots, run lm, summarize results, by species

```{r, echo=FALSE,results='hide', message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

summarise_all_fits <- function(x) {
  tmp <- lapply(x, summarise_fit) %>%
    bind_rows
  data.frame(species=names(x), tmp)
}
```

### 2015-04-30 results_RAschedules.Rmd
continuing to organize results

#RA schedules (vs. propH)
- these curves are particularly good for a number of species such as EPMI, COER, PELA
- they also show that for PILI that among individuals that reproduce, the "potential" RA decreases with increasing propH

@@

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

@@

#RA schedules (by size)
```{r, echo=FALSE,message=FALSE}
plot(RA~FinalWeight, SummaryInd, log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

data <- SummaryInd
data <- split(data, data$species)

@@

#RA vs RGR - 
similar to the plots of RA vs Growth Investment, but relationships stronger. For these plots, all but one species (PEPU) have a significantly negative slope
the results are more of less the same regardless of whether one includes or excluded individuals where RA=0
```{r, echo=FALSE,message=FALSE}
plot(RA~RGR, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

@@

>>>>

#RA schedules (vs. propH)  
- these curves are particularly good for a number of species such as EPMI, COER, PELA  
- they also show that for PILI that among individuals that reproduce, the "potential" RA decreases with increasing propH  
- when all data points are included, all species except PEPU, PHPH, and PULU show increasing RA with increasing height  
- when RA =0 is excluded, the following species still have a significant positive correlation: BOLE, COER, EPMI, GRBU, GRSP, HATE. LEES and PELA have a p-value between 0.05-0.10   
- PEPU has a negative correlation - matches Mark's suggestion that it is "hanging on"

 NOTE: Again, a linear line is not necessarily the most appropriate

@@

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

@@

#RA schedules (by size)
1. These plots have always been unsatisfying - with few exceptions (PULU, PEPU, LEES) plots against height are stronger. And both PILI and PEPU have very low R^2 by either method  
2. If Only individuals with RA>0 are included only COER, EPMI, GRBU, HATE and LEES still have a significant positive correlation - for the remaining species correlation only exists because of small, non-reproductive plants having a strong effect on the curve

```{r, echo=FALSE,message=FALSE}
plot(RA~FinalWeight, SummaryInd, log="x", col=col.spp[as.factor(species)],pch=16)
legend("topright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

data <- subset(SummaryInd)
data <- split(data, data$species)

@@

#RA vs RGR - 
1. this is another way to look at the growth-reproduction trade-off 
2. similar to the plots of RA vs Growth Investment, but relationships stronger. For these plots, all but one species (PEPU) have a significantly negative slope  
3. the results are more of less the same regardless of whether one includes or excluded individuals where RA=0  
4. PEPU an exception because so few individuals reproducing and those that are reproducing have intermediate RGR - neither the young plants nor the older plants just hanging on  
5. PELA doesn't have a relationship either, because so few reproducing individuals. But if the many individuals with RA=0 added back in, strong negative relationship for PELA

```{r, echo=FALSE,message=FALSE}
plot(RA~RGR, SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
legend("bottomright",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

@@

data <- subset(SummaryInd,RA>0&RGR>.04)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  asin(sqrt(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

@@

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)

### 2015-04-30 results_growth.Rmd
continuing to organize results

source("scripts/summary_dataframes.R")
```
#Growth data, general patterns

#Generic functions for most plotting below
```{r, echo=FALSE, message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

@@

```{r, echo=FALSE, message=FALSE}
yvar <- ""
xvar <- ""
data <- filter(SummaryInd,

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar)

summary(mod)
anova(mod)
```

```{r, echo=FALSE, message=FALSE}

```

>>>>

source("scripts/summary_dataframes.R")
```
#Growth data, general patterns
- proportion of maxH matters more for some species than others. 
1. The 4 "canopy" species all tend to have high correlations between "prop maxH" and growth and reproductive investment (both relative and absolute)  
2. The lower-growing species are more variable, with "prop maxH mattering more for some than others"

- For almost all species there is a strong negative relationship between RGR and RA - indeed, more reproduction = less growth  
1. Meanwhile there is the strong positive relationship between growth investment and reproductive investment, because bigger plants grow and reproduce more.  
2. It is good to have a big enough data set to capture both trends - and to capture them consistently across many species

- Many plants have a clear upturn in RA and corresponding downturn in RGR at a specific prop maxH  
1. e.g. see EPMI for distinct pattern

Generic functions for most plotting below
```{r, echo=FALSE, message=FALSE}
plot_yvar_vs_xvar <- function(data, yvar = "y", xvar = "x", ...) {
  Y <- data[[yvar]]
  X <- data[[xvar]]
  plot(Y ~ X, data=data,  cex.axis=.8, las=1, pch=16, ...)
}

@@

#Comparing growth between different measures and across age

###Weight vs. height
- No surprises here , but need to fit proper lines to the data
- Plants asymptote in weight as they reach their maximum height. This is true for all species, although some species, like BOLE and GRBU, have a much messier scatter than others  
- The follow on of course is to look at growth measures vs. height   
- NOTE: All heights have been scaled 0-1 based on the maximum height of each species across all of the sites. This has no effect on within species analyses, but lets data from all species to be plotted together

```{r, echo=FALSE, message=FALSE}
plot((FinalWeight~PropH), SummaryInd, col=col.spp[as.factor(species)],pch=16,log="y")
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
plot((FinalWeight~StartBasalDiamAv), SummaryInd, col=col.spp[as.factor(species)],pch=16,log="xy")

yvar <- "FinalWeight"
xvar <- "PropH"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,log="y",
                    main=spp, xlab=xvar, ylab=yvar)
  abline(lmfits[[spp]])
}
```

###Growth Investment versus proportion max H
- This turns out to be one of the most interesting sets of graphs  
- For the canopy species, PELA, HATE, PEPU, BAER, investment in growth plateaus (i.e. growth relative to size decreases) at ~40% maxH  
- For the shorter and shorter-lived species there is much less of a relationship, or a complete lack of relationship (i.e. BOLE, PHPH)  
- Quite interesting, there is a distinctly negative relationship for GRBU; its absolute growth investment actually declines as it increases in size  
- For many of the species with a good relationship, it seems that there are two breakpoints in the curve - wondering what these mean with respect to reproductive onset and RA

```{r, echo=FALSE, message=FALSE}
plot((GrowthInv~PropH), SummaryInd, log="y", col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(GrowthInv~PropH, SummaryInd, log="y", col=col.age[as.factor(age)],pch=16)
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")

yvar <- "GrowthInv"
xvar <- "PropH"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(log(x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, log="y",
                    main=spp, xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

###Proportion max H by RGR
- Ignoring a few outliers, RGR declines with increasing height is 12 species  
- Only in HEPU and PILI is there no significant trend  
- The species with a poorer correlation are mostly shorter ones that are presumably often broken off   
- These are continued ways to look at the question of how important is height (relative to the species maxH) in understanding its growth and reproduction

```{r, echo=FALSE, message=FALSE}
plot((RGR~PropH), SummaryInd, col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

yvar <- "RGR"
xvar <- "PropH"

data <- SummaryInd
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  (x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar,
                    main=spp, xlab=xvar, ylab=yvar)
  abline(lmfits[[spp]])
}
```

###PropH by age
- Confirming that species get taller as they get older  
- Obviously true for most species, but some (i.e. PILI, somewhat BOLE, HEPU) they reach their "maxH" within a few years and then the relationship is mostly flat. These are of course also the species which are shorter lived and have high levels of early reproduction

```{r, echo=FALSE, message=FALSE}
yvar <- "PropH"
xvar <- "age"

data <- subset(SummaryInd)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm((x[[yvar]]) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

results$r.squared <- round(results$r.squared,digits=3)
results$p.value <- round(results$p.value,digits=3)
results$a <- round(results$a,digits=2)
results$b <- round(results$b,digits=2)

results

for(spp in names(data)) {
  plot_yvar_vs_xvar(data[[spp]], yvar, xvar, log="x",
                    main=spp, xlab=xvar, ylab=yvar)

  abline(lmfits[[spp]])
}
```

###Growth Investment versus weight   
- showing that slope continues to be 1:1, but intercept decreases with time because more standing biomass
```{r, echo=FALSE, message=FALSE}
plot((GrowthInv~FinalWeight), SummaryInd, log="xy", col=col.spp[as.factor(species)],pch=16)
abline(0,1,lwd=2)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot(GrowthInv~FinalWeight, SummaryInd, log="xy", col=col.age[as.factor(age)],pch=16)
abline(0,1,lwd=2)
legend("topleft",legend=names(labels.age),col=col.age,pch=16, cex=.8,bty="n")
```

#Comparing growth of whole plant vs. leader

###Correlations between different growth measures
- Increased cross-sectional area of a leader and the entire plant are fairly correlated and uneffected by age (once not baby)

TO DO: species should be a random effect for these stats- I want to know the effect of age on the relationship once the effect of species has been accounted for

```{r, echo=FALSE,message=FALSE}
plot(change_shoot_area~change_basal_area,SummaryInd, log="xy",col=col.age[as.factor(age)], pch=16, cex=.8,xlab="change in basal stem area",ylab="change in shoot area")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1,bty="n")
mod <- lm(change_shoot_area~change_basal_area*age*species,SummaryInd)
anova(mod)
```

#Comparing growth rates of species with different trait values

### Species with larger seeds have higher growth rates  
- The same basic relationship exists across all ages, so combined all data here  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ seed_size,data=subset(SummarySppAge,age>.5),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="seed size (mg)",ylab="growth investment (mg per year)",main="Effect of seed size on growth rates")
legend("topleft",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ seed_size, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

### Species with greater maximum height have higher growth rates
- For younger plants there is no relationship with *SPECIES* overall maximum H  
- There is a positive correlation with the maximum height the species reaches at a given age -i.e. fast growing young plants are tall when young.  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age<3),log="xy",col=col.age[as.factor(age)],pch=16,cex=1.1,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==2.4), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")

plot(GrowthInv_mean ~ maxH_spp,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1.1,xlab="maximum height (mm)",ylab="growth investment (mg per year)",xlim=c(500,3200),main="Relationship between growth rates on maximum H")
legend("bottomright",col=col.age, labels.age, pch=16, cex=1,bty="n")
text(GrowthInv_mean ~ maxH_spp, data=subset(SummarySppAge,age==9), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
```

### Species with higher wood density have lower growth rates  
- Adult plants strongly show this pattern
- However there is no relationship in younger plants  
```{r, echo=FALSE,results='hide', message=FALSE}
plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age>3),log="xy",col=col.age3[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==5), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomleft",col=col.age, labels.age, pch=16, cex=1, bty="n")

plot(GrowthInv_mean ~ WD_mean,data=subset(SummarySppAge,age<3),log="xy",col=col.age[as.factor(age)],pch=16,cex=1,xlab="wood density",ylab="growth investment (mg per year)",xlim=c(0.57,0.9),main="Relationship between wood density and growth rates")
text(GrowthInv_mean ~ WD_mean, data=subset(SummarySppAge,age==2.4), labels=species, pos=2, offset=-2.4, cex=0.9,col="blue")
legend("bottomleft",col=col.age, labels.age, pch=16, cex=1, bty="n")
```

### 2015-04-30 results_leaf_lifespan.rmd
continuing to organize results

#Plotting LMA vs. age
older plants have marginally higher LMA, but the effect mostly disappears when the youngest plants are excluded
there is enormous variation in LMA across species

```{r, echo=FALSE,message=FALSE}
mod <- lm(LMA_mean~species*age, SummarySppAge)
summary(mod)
anova(mod)

@@

@@

>>>>

#LMA vs. age
- older plants have marginally higher LMA, but the effect mostly disappears when the youngest plants are excluded  
- there is enormous variation in LMA across species  
- also had plot of just branches from 2nd-last leaf cohort - no real difference, so using mean LMA of up to 3 leaf cohorts

plot(LMA_mean~age, SummarySppAge,log="xy", col=col.spp[as.factor(species)], pch=16)
mod <- lm(LMA_mean~species*age, SummarySppAge)
summary(mod)
anova(mod)

@@

```{r, echo=FALSE,results='hide', message=FALSE}
plot(LMA~age,data=subset(LMA,branch_age=="mid"),log="xy", xlab="age",ylab="LMA",col=col.spp[as.factor(species)],pch=16)
legend("bottomright",col=col.spp, labels.spp, pch=16, cex=1)
```

@@

#PropH vs LL_death
- Have just plotted overall data, because no relationships across or within species; only GRSP has significant negative relationship and BOLE has significant positive relationship
```{r, echo=FALSE,message=FALSE}
plot((LL_death~PropH), data=subset(SummaryInd,LL_death<10), col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")

plot((PropH~log(age)), data=subset(SummaryInd), col=col.spp[as.factor(species)],pch=16)
legend("topleft",legend=names(labels.spp),col=col.spp,pch=16, cex=.8,bty="n")
```
